---
title: "Estimating multi-period geometric portfolio returns"
output:
  html_document: 
    keep_md: TRUE
---

The convention among actuaries is to use multi-period geometric returns for the purpose of discounting liabilities.  So, the question is how, exactly, are they calculated.

Let's say we have a portfolio with the following properties on a 30 year horizon.


```{r}
asset_names=c("BbgAgg","LrgUSEq","SmlUSEq","EAFE","EAFESml","EM","PrivCredit","RE")
geom_returns=c(4,7.5,7.75,7.75,9.5,9.5,8,6.5)/100
volatility=c(6.03,17.5,21,21,28,13,9.29,15)/100
weights=c(10,17,5,16,4,8,20,20)/100
correlation=matrix(nrow=8,data=
                     c(100,0,0,0,0,0,0,10,
                       0,100,90,70,60,70,60,40,
                       0,90,100,60,65,75,65,40,
                       0,70,60,100,70,60,75,35,
                       0,60,65,70,100,45,80,30,
                       0,70,75,60,45,100,65,50,
                       0,60,65,75,80,65,100,40,
                       10,40,40,35,30,50,40,100))/100
names(geom_returns)=names(volatility)=names(weights)=asset_names
rownames(correlation)=asset_names
colnames(correlation)=asset_names
```

```{r}
geom_returns
volatility
weights
correlation
```

We now calculate a covariance matrix from this information.

```{r}
Dsqrt=0*correlation
diag(Dsqrt)=volatility
covariance=Dsqrt%*%correlation%*%Dsqrt
round(covariance,2)
```

The portfolio volatility is calculated as follows.

```{r}
portvol=sqrt(t(weights)%*%covariance%*%weights)
```

So, your portfolio volatility is `r round(portvol*100,2)`%.  This compares to the weighted average asset volatility of `r round(100*(volatility%*%weights),2)`%.

Now let's calculate the expected geometric return of the portfolio.  To do this, we calculate the weighted average arithmetic return of the portfolio which we then adjust for the portfolio volatility.  These calculations are performed using a 30 year horizon to be consistent with the return assumptions.   

First we create functions to do the conversion.  The formula to convert between arithmetic and geometric returns is from <http://www.treasury.govt.nz/publications/research-policy/wp/2003/03-28/twp03-28.pdf>

```{r}
# function to convert from geom return to arith return
# geo is the geometric return (may be a vector)
# vol is the volatility of the asset (may be a vecor)
# n is the number of years
geom_to_arith=function(geo,vol,n) {
  var=vol^2
  factor=(1+var/((1+geo)^2))^((1-n)/(2*n))
  arith=-1+(1+geo)/factor
  return(arith)
}
# function to convert from arith return to geom return
# arith is the arithmetic return (may be a vector)
# vol is the volatility of the asset (may be a vecor)
# n is the number of years
arith_to_geom=function(arith,vol,n) {
  var=vol^2
  factor=(1+var/((1+arith)^2))^((1-n)/(2*n))
  geo=-1+(1+arith)*factor
  return(geo)
}
```

Convert the geometric returns to arithmetic returns.

```{r}
n=30
arith_returns=geom_to_arith(geom_returns,volatility,n)
round(arith_returns,4)
```

Next we calculate the portfolio arithmetic return (which is the weighted average of the arithmetic returns).

```{r}
port_arith_return=arith_returns%*%weights
port_arith_return
```

Finally, we calculate the portfolio geometric return.

```{r}
port_geom_return=arith_to_geom(port_arith_return,portvol,n)
wavg_geom_return=geom_returns%*%weights
port_geom_return
```

So, the expected geometric return of the portfolio is `r round(port_geom_return*100,2)`%.  This is `r round(10000*(port_geom_return-wavg_geom_return))`bp greater than the weighted average geometric return of `r round(100*wavg_geom_return,2)`%.


